A Computational Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-03-28 paper/paper.tex
more

In artificial intelligence (AI), a popular abstract model for
multi-step decision processes in such an environment are so-called
\textit{Partially Observable Markov Decision Processes (POMDPs)}.
A POMDP models a decision process in which it is assumed that environment dynamics are determined by a Markov process,
but the agent cannot directly observe the underlying state. Instead, it tries to optimize a \textit{subjective} reward
signal which is tried alongside each transition in the environments state-space, and maintains a probability distribution over the set of possible
states, based on a set of observations and observation probabilities. This is a minimal amount of assumptions that can be made about an environment,
and is characteristic of so-called \textit{model-free RL}, the approach adapted by our work.
In contrast, \textit{model-based RL} starts off with some mechanistic assumptions about the dynamics of world.
These can be assumptions about the physical laws governing the agent's world or constraints on the state-space and transitions thereupon.
For example if a rat in a maze knows, for sure,
that staying still will produce no change in the environment, and in particular will not eventually lead to finding the food.
% A core property of human intelligence has been proposed underly to improve
% on an expected utility outcomes as a strategy for action choice in uncertain
% environments \citep{gershman2015computational}.
Our rat might represent such knowledge about the world as:
\begin{itemize}
\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a cell / chamber containing food.
\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
\item etc.
\end{itemize}

@@

\subsection{Computing optimal behavior}
By construction, maximizing the expectation in  \eqref{eq:qval} is intractable in general
\textbf{XXX coputationally or statistically?}
due to the enormous number of possible
\textbf{XXX action???}
trajectories the agent might need to
consider.
% \begin{mdframed}
%   \textbf{N.B.:}
%   If the Bellman-based computational ideas have a corresponding
%   neurobiological implementation, this provides a framework of how
%   optimal behavior could be computable in the human brain.
% \end{mdframed}
One popular simplification that removes
%this issue
\textbf{excess in complexity on the agent's horizon???}
is to only consider the sub-family of
\textit{deterministic policies}, which map each state to the best single best action which
should be taken at thin state, and therefore define functions from states to actions.

@@

\subsubsection{Interlude: zoom on the Hippocampus}
\paragraph{Does the hippocampus do Monte-Carlo sampling ?}
In RL, Monte-Carlo simulation can be used to update the agent's belief state
\citep{silver2010monte}. Such methods have a sample complexity that is determined only by the underlying difficulty of the POMDP, rather than the size of the state space or observation space,
which can be prohibitively large.
Monte-Carlo simulation provides a simple method for evaluating a state $s$.
They provide an effective mechanism both for tree search and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible.

@@

In our proposed RL-based view, \textit{inference} in the human brain reduces to generalization of
  policy and value adaptations from sampled experiences to
  successful action choices and reward predictions in future states.
  Plasticity in the DMN also arises naturally:
  If an agent behaving optimally in a given environment moves
  to novel, yet unexperienced environments, reward prediction errors will
  massively increase. This will lead to adaptation of the policy matrix
  and value function until the system converges to a
  new steady-state of optimal action decisions given certain states.

>>>>

In artificial intelligence (AI), a popular abstract model for
multi-step decision processes in such an environment are so-called
\textit{Partially Observable Markov Decision Processes (POMDPs)}.
A POMDP models a decision process in which it is assumed that environment dynamics are determined by a Markov process,
but the agent cannot directly observe the underlying state. Instead, it tries to optimize a \textit{subjective} reward
signal which is tried alongside each transition in the environments state-space, and maintains a probability distribution over the set of possible
states, based on a set of observations and observation probabilities. This is a minimal amount of assumptions that can be made about an environment,
and is characteristic of so-called \textit{model-free RL}, the approach adapted by our work.
\begin{mdframed}
\textbf{Model-free RL in the brain is plausible.}  Indeed, as was noted in \cite{dayan2008decision}, it has long been suggested in the literature that there
  is a rather direct mapping of model-free RL learning algorithms
onto the brain, with the neuromodulator dopamine serving
as a teaching signal to train values or policies by controlling
synaptic plasticity at targets such as the ventral and
dorsolateral striatum.
\end{mdframed}

In contrast, \textit{model-based RL} starts off with some mechanistic assumptions about the dynamics of world.
These can be assumptions about the physical laws governing the agent's world or constraints on the state-space and transitions thereupon.
For example if a rat in a maze knows, for sure,
that staying still will produce no change in the environment, and in particular will not eventually lead to finding the food.
% A core property of human intelligence has been proposed underly to improve
% on an expected utility outcomes as a strategy for action choice in uncertain
% environments \citep{gershman2015computational}.
Our rat might represent such knowledge about the world as:
\begin{itemize}
\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a cell / chamber containing food.
\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
\item etc.
\end{itemize}

@@

\subsection{Computing optimal behavior}
By construction, maximizing the expectation in  \eqref{eq:qval} is computationally intractable in general
due to the enormous number of possible action
trajectories the agent might need to
consider.
% \begin{mdframed}
%   \textbf{N.B.:}
%   If the Bellman-based computational ideas have a corresponding
%   neurobiological implementation, this provides a framework of how
%   optimal behavior could be computable in the human brain.
% \end{mdframed}
One popular simplification that removes
this issue is to only consider the sub-family of
\textit{deterministic policies}, which map each state to the best single best action which
should be taken at thin state, and therefore define functions from states to actions.

@@

\subsubsection{Interlude: zoom on the Hippocampus}
\paragraph{Does the hippocampus do Monte-Carlo sampling ?}
In RL, Monte-Carlo simulation can be used to update the agent's belief state
\citep{silver2010monte}. Such methods have a sample complexity that is determined only by the underlying difficulty of the POMDP, rather than the size of the state space or observation space,
which can be prohibitively large.
Monte-Carlo simulation provides a simple method for evaluating the value / fitness of a state.
They provide an effective mechanism both for tree search and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible.

@@

  In our proposed model-free RL-based view, \textit{inference} in the human brain reduces to
  generalization of
  policy and value adaptations from sampled experiences to
  successful action choices and reward predictions in future states.
  Plasticity in the DMN also arises naturally:
  If an agent behaving optimally in a given environment moves
  to novel, yet unexperienced environments, reward prediction errors will
  massively increase. This will lead to adaptation of the policy until the system converges to a 
  new steady-state of optimal action decisions given certain states.
